---
title: "Time Series Analysis"
type: "Final Project"
author: "Daniel Bustillo Mac Lean"
matriculation_number: "609728"
date: "`r`Sys.date()"
university: "Humboldt-Universität zu Berlin"
institute: "School of Business and Economics" 
chair: "Institute for Statistics and Econometrics"
advisor1: "Prof. Dr. Simone Maxand"
program: "Master of Science in Economics"
# chair2: "Chair of Econometrics"
# abstract: |  
#   `r if(knitr:::is_latex_output()) paste(readLines("prelim/00-abstract.Rmd"), collapse = '\n  ')`
# acknowledge: |  
#   `r if(knitr:::is_latex_output()) paste(readLines("prelim/00-acknowledge.Rmd"), collapse = '\n  ')`
# abbreviations: |  
#   `r if(knitr:::is_latex_output()) paste(readLines("prelim/00-abbreviations.Rmd"), collapse = '\n  ')`
# lot: true                              # list of tables
# lof: true                              # list of figures
# link-citations: true
# bibliography: bib/references.bib       # bibliography file
knit: "bookdown::render_book"
output: 
  huwiwidown::thesis_pdf  
space_between_paragraphs: true       # if you want spaces bt paragraph
# header-includes:     # if including additional latex packages or commands
  # - \setlength{\parindent}{17pt}
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of
metadata used to produce the document.  Be careful with spacing in this header!
If you'd like to include a comment that won't be produced in your resulting file
enclose it in a block like this.
-->
```
```{=html}
<!--
If you receive a duplicate label error after knitting, make sure to delete the
index.Rmd file and then knit again.
-->
```
```{r include_packages, include = FALSE}
# This chunk ensures that the `huwiwidown` package is installed and loaded. This
# package includes the template files for the thesis.
if (!require(devtools)) {
  install.packages("devtools", repos = "http://cran.rstudio.com")
}
if (!require(huwiwidown)) {
  devtools::install_github("phister/huwiwidown")
  library(huwiwidown)
}
```

```{=html}
<!-- 
The {.unnumbered} option here means that the introduction will be "Section  0."
You can also use {-} for no numbers on section. 
-->
```
```{=html}
<!--
The body of the thesis starts underneath the text below. Feel free to delete
everything below this.
-->
```
# Motivation

Increasing evidence has shown that human emissions of carbon dioxide and
other greenhouse gases are a primary driver of climate change (IPCC,
2013). This makes worldwide emissions one of the world's most pressing
challenges and has provoked various international agreement, like the
Paris Agreement on emission reduction and other climate goals. Three of
the largest current emitting regions are China, the US and the EU.

# Introduction

The main topic of this term paper is to analyze and model the trends of
CO2 emissions in some of the countries of the European Union, the United
States of America and China (and perhaps comparing them with emerging
economics such as Brazil and India). Current actions to mitigate the
climate effects of such steep rise of CO2 emissions over the last 50-60
years (citation needed) are not enough to reach the goals set in the
Paris Agreement by 2030. Such is the slowed response from the governing
institutions and international organizations that the current trends
indicate that the temperature increases (and all the consequences behind
it) will be irreversible in less that 5 years(citation needed). This
paper will analyze the trends of CO2 emissions of the above-mentioned
countries and will design a fitting econometric model. Lastly, a
forecast for the next periods based on the model will be presented.

Some exploratory data analysis on the different datasets

What kind of data do we use? C02 emissions per capita, why? Arguments
for using that measure -\> Although all 3 regions are really big, the
population density is different and thus "spreading" the CO2 emissions
by population gives us more sensible data and accounted for the
different sizes of population.

Plot of CO2 in Europe

Plot of CO2 in USA

Plot of CO2 in China

(Plot of CO2 in Brazil

Plot of CO2 in India)

The econometric and economic theory is briefly described and related
literature is presented

Idea: compare with GDP or GNP

Include kyoto and paris agreement and other historical events
(fukushima, fridays for future?)

# The model and data

The econometric model and the data applied are described, a first
descriptive analysis of the data is shown

-   Describe the data and its quality.All data is taken from the World
    Bank's Database, overall this database has proven to be reliable.

-   How was the data sample selected?

    As stated in the Introduction, our interest here is to analyze the
    trends of the 3 biggest emitters of Co2 in the world: the USA, the
    European Union and China [quote on that]. Additionally, we compare
    those trends and forecast, with trends and forecast of emerging
    economies such as Brazil and India, as the

-   Provide descriptive statistics such as:

    -   time period,

    -   item number of observations, data frequency,

    -   item mean, median,

    -   item min, max, standard deviation,

    -   item skewness, kurtosis, Jarque--Bera statistic,

    -   item time series plots, histogram.

# Estimation and results

The model choice is reasoned well in statistical and economic terms, the
estimation results are presented, the results are interpreted and
connected to the economic theory and literature from point 1.

Find an appropriate ARIMA model to describe the three series. Base your
decision on (P)ACFs, unit root tests, AIC and further diagnostic checks.
What limits the validity of these results? 4. Forecast the CO2 emissions
one period ahead and for 2030 based on the chosen model. Evaluate the
forecast in comparison to an alternative prediction method. 5. Discuss
the economic consequences of your results and suggest different
techniques how your model and a potential long-term forecast could be
improved.

# Conclusion:

Summary of the main results and outlook on further interesting research
directions

<!--chapter:end:index.Rmd-->

# Introduction

-   What is the subject of the study? Describe the economic/econometric
    problem.

-   What is the purpose of the study (working hypothesis)?

-   What do we already know about the subject (literature review)? Use
    citations: @lingenfelser2011systematic shows that... Alternative
    Forms of the Wald test are considered [@kuncheva2004combining].

-   What is the innovation of the study?

-   Provide an overview of your results.

-   Outline of the paper:\
    *The paper is organized as follows. The next section describes the
    model under investigation. Section "[Data]" describes the data set
    and Section "[Results]" presents the results. Finally, Section
    "[Conclusion](#conclusion-1)" concludes.*

-   The introduction should not be longer than 4 pages.

<!--chapter:end:sections/01-intro.Rmd-->

# Methodology

-   How was the data analyzed ?

-   Present the underlying economic model/theory and give reasons why it
    is suitable to answer the given problem[^1].

-   Present econometric/statistical estimation method and give reasons
    why it is suitable to answer the given problem.

-   Allows the reader to judge the validity of the study and its
    findings.

-   Depending on the topic this section can also be split up into
    separate sections.

[^1]: Here is an example of a footnote.

<!--chapter:end:sections/02-method.Rmd-->

# Data

-   Describe the data and its quality.

-   How was the data sample selected?

-   Provide descriptive statistics such as:

    -   time period,

    -   item number of observations, data frequency,

    -   item mean, median,

    -   item min, max, standard deviation,

    -   item skewness, kurtosis, Jarque--Bera statistic,

    -   item time series plots, histogram.

-   For example:

```{r table1, echo=FALSE}
col_names <- c(
  "3m", "6m", "1yr", "2yr", "3yr", "5yr", "7yr", "10yr", "12yr", "15yr"
)
means     <- c(
  3.138, 3.191, 3.307, 3.544, 3.756, 4.093, 4.354, 4.621, 4.741, 4.878
)
stddev    <- c(
  0.915, 0.919, 0.935, 0.910, 0.876, 0.825, 0.803, 0.776, 0.768, 0.762
)
row_names <- c("Mean", "StD")

df <- matrix(data = c(means, stddev), nrow = 2, byrow = T)
rownames(df) <- row_names
colnames(df) <- col_names
df <- data.frame(df)
knitr::kable(
  df,
  booktabs = TRUE,
  caption = "Detailed descriptive statistics of location and dispersion for 2100 observed swap rates for the period from February 15, 1999 to March 2, 2007. Swap rates measured as 3.12 (instead of 0.0312).",
  col.names = col_names, 
  escape = FALSE
) %>%
kable_styling(latex_options = c("HOLD_position")) %>%
row_spec(2, hline_after = T)
```

-   Allows the reader to judge whether the sample is biased or to
    evaluate possible impacts of outliers, for example.

-   Here tables can be easily integrated using the `kable()` function in
    the `knitr` package (with perhaps some additional help from the
    `kableExtra` package). `kable()` will automatically generate a label
    for the table environment. That way you don't have to manually enter
    in the table in LaTex, you can embed tables from R code.

-   Tables can be referenced using `\@ref(label)`, where `label` is
    `tab:<name>`, where `<name>` is the code chunk label.

-   The appearance may look different to tables directly typed with
    LaTex, due to limitations in `kable()`. To compare:

    ```{=tex}
    \begin{table}[ht]

      \begin{center}
          {\footnotesize
          \begin{tabular}{l|cccccccccc}
              \hline \hline
                        & 3m    & 6m    & 1yr   & 2yr   & 3yr   & 5yr   & 7yr   & 10yr  & 12yr  & 15yr   \\
              \hline
                  Mean   & 3.138 & 3.191 & 3.307 & 3.544 & 3.756 & 4.093 & 4.354 & 4.621 & 4.741 & 4.878  \\
                  StD    & 0.915 & 0.919 & 0.935 & 0.910 & 0.876 & 0.825 & 0.803 & 0.776 & 0.768 & 0.762  \\
              \hline \hline
          \end{tabular}}
      \end{center}
      \caption{This table was handwritten with LaTeX.}
      \label{tab:table2}
      \end{table}
    ```

<!--chapter:end:sections/03-data.Rmd-->

# Results

-   Organize material and present results.

-   Use tables, figures (but prefer visual presentation):

    -   Tables and figures should supplement (and not duplicate) the
        text.

    -   Tables and figures should be provided with legends.

    -   *Figure* \@ref(fig:graph) *shows how to include and reference
        graphics. The graphic must be labelled before. Files must be in
        **.eps** format. You can do this really easily in R Markdown
        with `knitr::include_graphics()`*!

    -   Figures can be referenced with `\@ref(fig:<name>)`, where
        `<name>` is the name of the code chunk.

    -   Tables and graphics may appear in the text or in the appendix,
        especially if there are many simulation results tabulated, but
        is also depends on the study and number of tables resp. figures.
        The key graphs and tables must appear in the text!

-   R Markdown can also supports math equations just like *LaTeX*!

    -   *Equation* \@ref(eq:SpecDens) *represents the ACs of a
        stationary stochastic process:* \begin{equation}
                f_y(\lambda) = (2\pi)^{-1} \sum_{j=-\infty}^{\infty}
                               \gamma_j e^{-i\lambda j}
                             =(2\pi)^{-1}\left(\gamma_0 + 2 \sum_{j=1}^{\infty}
            \gamma_j \cos(\lambda j)\right)
                                           (\#eq:SpecDens)
        \end{equation} *where* $i=\sqrt{-1}$ is the imaginary unit,
        $\lambda \in [-\pi, \pi]$ is the frequency and the $\gamma_j$
        are the autocovariances of $y_t$.

    -   Equations can be referenced with `\@ref(eq:<name>)`, where name
        is defined by adding `(\#eq:<name>)` in the line immediately
        before `\end{equation}`.

## Review of Results

-   Do the results support or do they contradict economic theory ?

-   What does the reader learn from the results?

-   Try to give an intuition for your results.

-   Provide robustness checks.

-   Compare to previous research.

<!--chapter:end:sections/04-results.Rmd-->

# Conclusion {#conclusion-1}

-   Give a short summary of what has been done and what has been found.

-   Expose results concisely.

-   Draw conclusions about the problem studied. What are the
    implications of your findings?

-   Point out some limitations of study (assist reader in judging
    validity of findings).

-   Suggest issues for future research.

<!--chapter:end:sections/05-conclusion.Rmd-->

\newpage

# References {.unnumbered}

```{=html}
<!--
To remove the indentation of the first entry.
-->
```
\noindent

```{=html}
<!--
To create a hanging indent and spacing between entries.  These three lines may need to be removed for styles that don't require the hanging indent.
-->
```
```{=tex}
\setlength{\parindent}{-0.5cm}
\setlength{\leftskip}{0.5cm}
\setlength{\parskip}{8pt}
```
::: {#refs}
:::

```{=html}
<!--
Restore to defaults
-->
```
```{=tex}
\indent
\setlength{\parindent}{17pt}
\setlength{\leftskip}{0pt}
\setlength{\parskip}{0pt}
```
\newpage

<!--chapter:end:sections/98-references.Rmd-->

\appendix

# Appendix

Here goes the appendix!

## Figures

## Tables

```{=tex}
\begin{table}[ht]
    \begin{center}
        {\footnotesize
        \begin{tabular}{l|cccccccccc}
        \hline \hline
                        & 3m    & 6m    & 1yr   & 2yr   & 3yr   & 5yr   & 7yr   & 10yr  & 12yr  & 15yr   \\
            \hline
                Mean   & 3.138 & 3.191 & 3.307 & 3.544 & 3.756 & 4.093 & 4.354 & 4.621 & 4.741 & 4.878  \\
                Median & 3.013 & 3.109 & 3.228 & 3.490 & 3.680 & 3.906 & 4.117 & 4.420 & 4.575 & 4.759  \\
                Min    & 1.984 & 1.950 & 1.956 & 2.010 & 2.240 & 2.615 & 2.850 & 3.120 & 3.250 & 3.395  \\
                Max    & 5.211 & 5.274 & 5.415 & 5.583 & 5.698 & 5.805 & 5.900 & 6.031 & 6.150 & 6.295  \\
                StD    & 0.915 & 0.919 & 0.935 & 0.910 & 0.876 & 0.825 & 0.803 & 0.776 & 0.768 & 0.762  \\
            \hline \hline
        \end{tabular}}
    \end{center}
    \caption{Detailed descriptive statistics of location and dispersion for
    2100 observed swap rates for the period from
    February 15, 1999 to March 2, 2007. Swap rates measured as 3.12 (instead of 0.0312).}
    \label{tab:apptable}
\end{table}
```
\newpage

<!--chapter:end:sections/99-appendix.Rmd-->
