---
title: Forecasting CO2 emissions using ARIMA models in Brazil, China, EU, India and
  US
author: "Daniel Bustillo Mac Lean"
date: "31-08-2021"
output:
  pdf_document:
    keep_tex: yes
    toc: yes
    toc_depth: 2
    latex_engine: xelatex
    citation_package: natbib
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_html: default
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
subtitle: Time Series Analysis Term Paper
link-citations: yes
editor_options:
  markdown:
    wrap: 72
bibliography: references.bib
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'),echo = FALSE)
options(htmltools.dir.version = FALSE)
```

```{r load libraries and datasets, include=FALSE}
library(tidyverse)
library(forecast)
library(tseries)
library(data.table)
library(reshape2)
library(zoo)
library(scales)
library(knitr)
library(Hmisc)
library(urca)
library(kableExtra)

data <- read.csv("Dataset/API_EN.ATM.CO2E.PC_DS2_en_csv_v2_2708833.csv", skip = 3, header=TRUE)
data <- data[,-c(2:4)]
```

# Motivation

Increasing evidence has shown that human emissions of carbon dioxide and
other greenhouse gases are a primary driver of climate change (@stocker2014climate). This makes worldwide emissions one of the world's most pressing
challenges and has provoked various international agreement, like the
Paris Agreement on emission reduction and other climate goals. Three of
the largest current emitting regions are China, the US and the EU.

\newpage

# Introduction

The main topic of this term paper is to analyze and model the trends of
CO2 emissions in some of the countries of the European Union, the United
States of America and China (and perhaps comparing them with emerging
economics such as Brazil and India). Current actions to mitigate the
climate effects of such steep rise of CO2 emissions over the last 50-60
years are not enough to reach the goals set in the
Paris Agreement by 2030. Such is the slow response from the governing
institutions and international organizations that the current trends
indicate that the temperature increases (and all the consequences behind
it) will be irreversible in the future. This
paper will analyze the trends of CO2 emissions of the above-mentioned
countries and will design a fitting Auto Regressive Integrated Moving
Average (ARIMA) model. Next, we will run diagnostics and perform the
necessary tests to assure that the model accounts for stationarity and
possibly seasonality. Lastly, a forecast for the next periods based on
the model will be presented.

# Literature review

There have been several studies that apply ARIMA models and similar
statistical and econometric techniques to forecast carbon dioxide
emissions for different regions, countries and time periods.

@fatima2019forecasting used Simple Exponential Smoothing (SES) and ARIMA
models to forecast CO2 emissions for several Asian countries, with China
and India among them. For China they fit an (1,2,0) ARIMA model and for
India an (0,2,1) ARIMA model.

@nyoni2019modeling modeled an ARIMA model for CO2 emissions in China for
the period 1960-2014, they found that an ARIMA(1,2,1) model is the most
suitable model to forecast total annual CO2 emissions for China.

A study on carbon dioxide emissions between 1972 and 2015 in Bangladesh
was conducted by @rahman2017modeling. According to their results, the
best fitting ARIMA model was of order (0,2,1).

One of the justifications some of these authors mention when choosing to
forecast using ARIMA models was that no other data is necessary, the
forecasting can be done using only historical values of the data, with
no other variables involved.

# The model and data

In this section, we go over the data, its characteristics, sources and
reliability. Moreover, we introduce the ARIMA (Auto Regressive
Integrated Moving Average) model in a formal way, next, we run the model
on our data to then set up all the insights to analyze in the following
section.

All data is taken from the World Bank's Database. This database has
proven to be reliable for industrialized countries, not so much for
developing countries, since data from these countries might not be
recollected or administered correctly by the corresponding authorities.

As stated in the Introduction, our interest here is to analyze the
trends of the 3 biggest emitters of CO2 in the world: the United States
of America, the European Union and China . Additionally, we compare
those trends and forecast, with trends and forecast of emerging
economies such as Brazil and India, as those countries display worrying
trends on CO2 emissions that are product of contamination and
deforestation due to their growing industries.

The measure we use for our analysis is C02 emissions per capita, because
although all country groups are the among the biggest in terms of
surface, the population density is different and thus "spreading" the
CO2 emissions by population gives us more sensible data and accounted
for the different sizes of population.

```{r}
#First filter the countries we are interested in
data <- data %>% 
  filter(Country.Name %in% c("European Union", "United States", "China", "India", "Brazil"))
names <- data$Country.Name
#transpose the dataframe to a u(seful format
data1 <- transpose(data)
colnames(data1) <- names
rownames(data1) <- colnames(data)
data1 <- data1[-1,]
#Clean rownames and remove last 3 empty entries
rownames(data1) <- lapply(rownames(data1), function(x) gsub(pattern = "X", replacement = "",x))
data1 <- tibble::rownames_to_column(data1,var = "Year")
data1 <- data1[-c(60:62),]
data1$Year <- as.Date(data1$Year,format = "%Y")

###Reorder code!!!
data1$Brazil <- as.numeric(data1$Brazil)
data1$China <- as.numeric(data1$China)
data1$`European Union` <- as.numeric(data1$`European Union`)
data1$India <- as.numeric(data1$India)
data1$`United States` <- as.numeric(data1$`United States`)
#melt
datap <- melt(data = data1,id="Year")
##Plot
ggplot(datap)+ geom_line(aes(x=Year, y= value, group=variable, colour= variable))+
   theme_classic() +
  xlab(label = "Year")+
  ylab("CO2 Emissions (metric tons per capita)")+
  ggtitle(label = "CO2 Emissions", subtitle = "in metric tons per capita from 1960 to 2018")+
  theme(legend.title = element_text())


  
  


```

As seen above, CO2 emissions of the countries of interest are showing
different trends and magnitudes.[^1]

[^1]: For a more detailed view of each countries' time series see next
    section of the paper

Even though the European Union and the United States show some decline
in emissions after the 1980s, the level of emissions are significantly
higher than emerging countries like Brazil or India. The worrying part
is that China's carbon dioxide emissions surpassed the ones from the
European Union in the last 10 years, the steep rise since the early
2000s matches Chinas rising role as a main actor in the global economy.

Below are the summary statistics for all countries and their respective
time series plots (the time period of the data recollected is from 1960
to 2018 for all countries)

```{r  sumstats}
##Tables with descr stats
summary(data1[,names]) %>% kable(digits = 2, caption = "Summary statistics for all countries") %>% 
  kableExtra::kable_classic()

```

\newpage

### Individual Time series plots

To have a more detailed view of each countries dioxide emissions we plot
each time series plot individually.

```{r  descrstats,fig.show="hold", out.width="50%"}

## Create time series objects for oll countries
br <- ts(data1$Brazil,start = "1960")
autoplot(as.zoo(br), geom = "line") + scale_x_yearqtr(format = "%Y")+
  theme_classic()+
  xlab("Year")+
  ylab("CO2 Emissions (metric tons per capita)")+
  ggtitle("Brazil")

ch <- ts(data1$China,start = "1960")
autoplot(as.zoo(ch), geom = "line") + scale_x_yearqtr(format = "%Y")+
  theme_classic()+
  xlab("Year")+
  ylab("CO2 Emissions (metric tons per capita)")+
  ggtitle("China")
eur <- ts(data1$`European Union`, start = "1960")
autoplot(as.zoo(eur), geom = "line") + scale_x_yearqtr(format = "%Y")+
  theme_classic()+
  xlab("Year")+
  ylab("CO2 Emissions (metric tons per capita)")+
  ggtitle("European Union")

ind <- ts(data1$India, start="1960")
autoplot(as.zoo(ind), geom = "line") + scale_x_yearqtr(format = "%Y")+
  theme_classic()+
  xlab("Year")+
  ylab("CO2 Emissions (metric tons per capita)")+
  ggtitle("India")

us <- ts(data1$`United States`, start="1960")
autoplot(as.zoo(us), geom = "line") + scale_x_yearqtr(format = "%Y")+
  theme_classic()+
  xlab("Year")+
  ylab("CO2 Emissions (metric tons per capita)")+
  ggtitle("United States")
```

Plotting the time series individually makes it evident that all series
aren't stationary, as they show different trends. In the case of Brazil,
China and India, their emissions have been rising over the last 50 years
and there is positive trend. In the EU and the US, the case is
different, but they appear to be non-stationary.

\newpage

### Autocorrelation plots and Partial Autocorrelation plots

Using the autocorrelation function formula: $$\rho_k =
\frac{\gamma_k}{\gamma_o}=\frac{Cov[y_t,y_{t-k}]}{Var(y_t)}$$ all ACF
are calculated for each country group. This works as a visual way to
determine if a time series object is stationary or not. In the next
section we run the necessary test to confirm our findings from the ACFS.

```{r acfs_plots, fig.show="hold", out.width="50%"}
# Plot all acfs 
acf(br, main="Brazil")

acf(ch, main="China")

acf(eur,main="European Union")

acf(ind,main="India")

acf(us,main="United States")
```

To further assist with our model choice, we use the Partial
Autocorrelation Functions plots as well.

```{r pacf, fig.show="hold", out.width="50%"}
### PACF
pacf(br,main="Brazil")

pacf(ch,main="China")

pacf(eur,main="European Union")

pacf(ind,main="India")

pacf(us,main="United States")

```

\newpage

# Stationarity tests

### Augmented Dickey Fuller Test

With this test we want to negate the null hypothesis that a unit root is
present in an auto regressive model of a given time series, and that the
process is thus not stationary.

A common rule of thumb for determining pmax, suggested by
@schwert1989business, is $$ p_{max}=[12(\frac{T}{100})^{1/4}]$$

```{r adf}
# Iterate for each country

pmax = round(12*(length(br)/100)^0.25, 0)  # truncation lag according to Schwert (1989)
#Brazil 
br.adf = ur.df(br, type="trend", selectlags="AIC", lags=pmax)
br.adf_tp = punitroot(q=br.adf@teststat[1], N=length(br), trend="c", statistic="t") # approximate p-values by MacKinnon (1996) 
#China
ch.adf = ur.df(ch, type="trend", selectlags="AIC", lags=pmax)
ch.adf_tp = punitroot(q=ch.adf@teststat[1], N=length(ch), trend="c", statistic="t")
#EU
eur.adf = ur.df(eur, type="trend", selectlags="AIC", lags=pmax)
eur.adf_tp = punitroot(q=eur.adf@teststat[1], N=length(eur), trend="c", statistic="t")
#India
ind.adf = ur.df(ind, type="trend", selectlags="AIC", lags=pmax)
ind.adf_tp = punitroot(q=ind.adf@teststat[1], N=length(ind), trend="c", statistic="t")
#US
us.adf = ur.df(us, type="trend", selectlags="AIC", lags=pmax)
us.adf_tp = punitroot(q=us.adf@teststat[1], N=length(us), trend="c", statistic="t")

```

### KPSS Unit Root Test

We also use the KPSS Unit Root test from @kwiatkowski1992testing to test
the H0 hypothesis of stationarity.

```{r kpss, warning=FALSE}
br.kpss = kpss.test(br, null="Level", lshort=FALSE)
ch.kpss = kpss.test(ch, null="Level", lshort=FALSE)
eur.kpss = kpss.test(eur, null="Level", lshort=FALSE)
ind.kpss = kpss.test(ind, null="Level", lshort=FALSE)
us.kpss = kpss.test(us, null="Level", lshort=FALSE)

br.kpss.p <- br.kpss$p.value
ch.kpss.p <- ch.kpss$p.value
eur.kpss.p <- eur.kpss$p.value
ind.kpss.p <- ind.kpss$p.value
us.kpss.p <- us.kpss$p.value

kpss <- c(br.kpss.p, ch.kpss.p, eur.kpss.p,ind.kpss.p, us.kpss.p)
# br %>% ur.kpss() %>% summary()
# ch %>% ur.kpss() %>% summary()
# eur %>% ur.kpss() %>% summary()
# ind %>% ur.kpss() %>% summary()
# us %>% ur.kpss() %>% summary()
```

```{r tabunit}

unit <- data.frame(row.names = names, c(br.adf_tp, ch.adf_tp,eur.adf_tp,ind.adf_tp, us.adf_tp), kpss)
colnames(unit) <- c(" ADF p-value", "KPSS p-value")
unit %>% 
  kable(caption = "p-values for ADF and KPSS Tests") %>% 
  kableExtra::kable_classic() %>%
  kable_styling(latex_options = "HOLD_position")


```

Based on the table above , results from the Augmented Dickey Fuller Test
and the KPSS Test, we can only reject the ADF H0 hypothesis of
non-stationarity for Brazil and the European Union. On the contrary,
KPSS p-values show that we might reject the H0 hypothesis of
stationarity for Brazil, China and India.

\newpage

# Estimation and results

Having conducted the necessary test and evaluated al the different time
series and their characteristics, it is time to estimate a model,
present, analyze and interpret the results.

For the estimation of the ARIMA model we ran a so-called
[Grid-Search](https://www.r-bloggers.com/2018/11/searching-for-the-optimal-hyper-parameters-of-an-arima-model-in-parallel-the-tidy-gridsearch-approach/)
for the 5 different country groups and all possible combinations of
models within our parameters. We then evaluate the best model for each
country based on the Akaike Information Criterion :
$$ AIC=log(\hat{\sigma}^2_{\epsilon}) + 2\frac{p+q+1}{T}$$ Having
defined the maximal *p* order in the previous section (11 lags), we
restricted the *d* order to 3, as too many differentiation of the time
series don't bring better results. Moreover, we restrict the *q*
parameter from the MA(q) model to 11.

Below there is a table with each countries' best models and their
respective AIC coefficients.

```{r gridsearch ,eval=FALSE}
#Warning! This code chunk runs 576 models (combinations of (p,d,q) arima models)
#it takes a lot of time
order_list = list(seq(0, 11),
                   seq(0, 3),
                   seq(0, 11)) %>%
    cross() %>%
    purrr::map(lift_dl(c))
orderdf = tibble("order" = order_list)

models_df.br = orderdf %>%
    mutate(models = map(order, ~possibly(arima, otherwise = NULL)(x = br, order = .x))) %>% 
    filter(models != 'NULL') %>% 
    mutate(aic = map_dbl(models, "aic"))
best_model.br = models_df.br %>%
    filter(aic == min(models_df.br$aic, na.rm = TRUE))

models_df.ch = orderdf %>%
    mutate(models = map(order, ~possibly(arima, otherwise = NULL)(x = ch, order = .x))) %>% 
    filter(models != 'NULL') %>% 
    mutate(aic = map_dbl(models, "aic"))
best_model.ch = models_df.ch %>%
    filter(aic == min(models_df.ch$aic, na.rm = TRUE))

models_df.eur = orderdf %>%
    mutate(models = map(order, ~possibly(arima, otherwise = NULL)(x = eur, order = .x))) %>% 
    filter(models != 'NULL') %>% 
    mutate(aic = map_dbl(models, "aic"))
best_model.eur = models_df.eur %>%
    filter(aic == min(models_df.eur$aic, na.rm = TRUE))

models_df.ind = orderdf %>%
    mutate(models = map(order, ~possibly(arima, otherwise = NULL)(x = ind, order = .x))) %>% 
    filter(models != 'NULL') %>% 
    mutate(aic = map_dbl(models, "aic"))
best_model.ind = models_df.ind %>%
    filter(aic == min(models_df.ind$aic, na.rm = TRUE))

models_df.us = orderdf %>%
    mutate(models = map(order, ~possibly(arima, otherwise = NULL)(x = us, order = .x))) %>% 
    filter(models != 'NULL') %>% 
    mutate(aic = map_dbl(models, "aic"))
best_model.us = models_df.us %>%
    filter(aic == min(models_df.us$aic, na.rm = TRUE))
```

```{r table}

# In order to store the values from the past chunk and speed up the process
# we store the values in a data frame manually
results <- tibble("Country"=names,"Order"=c("(2,1,5)","(1,1,0)","(0,2,1)","(1,1,2)","(0,1,1)"), "AIC"=c("-127.6029","-61.7693","10.93498","-254.3758","98.9266"))

results %>% kable() %>% 
  kableExtra::kable_classic()%>%
  kable_styling(latex_options = "HOLD_position")


```

```{r arima2}
# sanity check, compare auto arima models vs the ones generated by Grid Search

br.autoarima <- auto.arima(y = br, d=2,stepwise = TRUE ,seasonal = FALSE, ic = "aic", test='adf')#this doesn't result in the past value

br.arima <- Arima(br,order = c(2,1,5))

ch.autoarima <- auto.arima(y=ch, max.p = pmax,max.q = pmax,max.d = 1,seasonal = FALSE,ic = "aic")

eur.autoarima <- auto.arima(y=eur, max.p = pmax,max.q = pmax,max.d = 2,seasonal = FALSE,ic = "aic")

ind.autoarima <- auto.arima(y=ind, max.p = pmax,max.q = pmax,d=1,seasonal = FALSE,ic = "aic", test = "kpss")

us.autoarima <- auto.arima(y=us, max.p = pmax,max.q = pmax, max.d = 1,seasonal = FALSE,ic = "aic", test="kpss")

```

We compared the results with results from the **auto.arima** function
from the *forecast* package, and they match for all the countries,
although the **auto.arima** function required some extra tweaking. In
the end, we choose the models that display the lowest AIC values.

### Diagnostics

Next we proceed to run diagnostics of our models, that means that our model should capture the dynamics of the time series. Consequently, the residuals should be approximately white noise (No residual correlation, same variance, normally distributed). We do this with the Ljung-Box Test: $$ Q_K= T( T+2) \sum^K_{k=1}\frac{1}{T-k}\hat{\rho}^2_k  \longrightarrow \chi^2 (K -p-q)  $$


```{r resid, fig.show="hold", out.width="50%", warning=FALSE}

# DIAGNOSTIC CHECKS BR
p <- br.arima$arma[1]
q <- br.arima$arma[2]

br.e <- br.arima$residuals    # COMPUTE RESIDUALS
plot (br.e, type="l")
Acf(br.e)
LB.br.e <- Box.test(br.e, lag=pmax, type="Ljung-Box", fitdf=(p+q))    # LJUNG-BOX TEST



# DIAGNOSTIC CHECKS CH
p <- ch.autoarima$arma[1]
q <- ch.autoarima$arma[2]

ch.e <- ch.autoarima$residuals    # COMPUTE RESIDUALS
plot (ch.e, type="l")
Acf(ch.e,11)
LB.ch.e <- Box.test(ch.e, lag=pmax, type="Ljung-Box", fitdf=(p+q))    # LJUNG-BOX TEST

# DIAGNOSTIC CHECKS EUR
p <- eur.autoarima$arma[1]
q <- eur.autoarima$arma[2]

eur.e <- eur.autoarima$residuals    # COMPUTE RESIDUALS
plot (eur.e, type="l")
Acf(eur.e,11)
LB.eur.e <- Box.test(eur.e, lag=pmax, type="Ljung-Box", fitdf=(p+q))    # LJUNG-BOX TEST

# DIAGNOSTIC CHECKS IND
p <- ind.autoarima$arma[1]
q <- ind.autoarima$arma[2]

ind.e <- ind.autoarima$residuals    # COMPUTE RESIDUALS
plot (ind.e, type="l")
Acf(ind.e,11)
LB.ind.e <- Box.test(ind.e, lag=pmax, type="Ljung-Box", fitdf=(p+q))    # LJUNG-BOX TEST

# DIAGNOSTIC CHECKS 
p <- br.arima$arma[1]
q <- br.arima$arma[2]

br.e <- br.arima$residuals    # COMPUTE RESIDUALS
plot (br.e, type="l")
Acf(br.e,11)
LB.br.e <- Box.test(br.e, lag=, type="Ljung-Box", fitdf=(p+q))    # LJUNG-BOX TEST

```

# Forecast 
Finally, we proceed to forecast carbon dioxide emissions for each country group until 2030 based on each fitted ARIMA model. 



```{r forecast,fig.show="hold", out.width="50%"}
h <- 12      # FORECAST HORIZON 
#BR

f <- forecast(br.arima, h=h, level=0.9)
plot(f, main="Forecast for Brazil")
# CH
f <- forecast(ch.autoarima, h=h, level=0.9)
plot(f, main="Forecast for China")

# EUR
f <- forecast(eur.autoarima, h=h, level=0.9)
plot(f, main="Forecast for EU")

# IND
f <- forecast(ind.autoarima, h=h, level=0.9)
plot(f, main="Forecast for India")

# US
f <- forecast(us.autoarima, h=h, level=0.9)
plot(f, main="Forecast for US")
```
The forecast for the different countries show somewhat different outcomes, in the cases of  Brazil and the United States, prediction for the next 12 periods shows that the emissions will remain stable (within a 99% confidence interval).
For the cases of China and India, the forecasted values show some worrying trends, by 2030 both of these countries will still show rising CO2 emissions, although China's trend seems to slow down a bit compared to the early 2000's. 
Luckily, the forecast for the European Union shows that by 2030, their CO2 emissions will resemble the ones from the 1960s.


\newpage

# Conclusion

To wrap up this paper we will review the process of the modelling, the
estimation, the diagnostics and the forecasting of the different time
series.

Summing up what has been done, we compared time series from 5 different countries or country groups, to analyze and model the CO2 emissions using ARIMA models. Each country shows a different trend, magnitude and volatily in the data, so different parameters and different models based on assumptions had to be fitted for each country.
We used a variety of a Grid Search algorithm to navigate through different combinations of ARIMA (p,d,q) parameters as a way to support our initial stationarity tests and Auto Correlation Functions. In the end we chose each model based on the lowest AIC value for each country. 
Although literature on CO2 emissions is not particularly scarce, not many authors and researchers seem to use ARIMA models for this kind of forecasting. Probably there might exist another advanced techniques that better forecast CO2 emissions, for example the novel library *prophet* by Facebook, a collection of forecasting algorithms.

From an environmental point of view, the results and forecast of this study are negative, though United States and the EU have shown decreasing trends in their emissions, they are still the two biggest actors in the world. With China's surprisingly fast emergence in the global economy, the forecast for them shows the most worrying trend. China has already surpassed the EU's emissions and though they have somewhat slowed down the pace, they still show a rising trend. The main reason for including two other emerging economies like Brazil and India, was to check if China's disturbing trends also translate into these countries. Sadly for our environment, the answer is yes, albeit on a much smaller scale.


Some other factors that could prove to be pivotal in the slow decline of
CO2 emissions in the European Union and USA could be historical events
like the Kyoto Protocol in 1997 and the Paris Agreement in 2015, and some
natural distasters, like the 2011 tsunami in Japan that caused the
explosion in the nuclear factory in Fukushima. This event led to a chain
of events that ended with Germany's (and some other European countries)
abandonment of atomic energy.


\newpage

# Appendix

For sake of completeness, we include here the differenced plots of the
time series and their respective ACFs.

```{r  diffplots,fig.show="hold", out.width="50%"}

## Create time series objects for oll countries
br <- ts(data1$Brazil,start = "1960")
autoplot(as.zoo(diff(br)), geom = "line") + scale_x_yearqtr(format = "%Y")+
  theme_classic()+
  xlab("Year")+
  ylab("CO2 Emissions (metric tons per capita)")+
  ggtitle("Brazil")

ch <- ts(data1$China,start = "1960")
autoplot(as.zoo(diff(ch)), geom = "line") + scale_x_yearqtr(format = "%Y")+
  theme_classic()+
  xlab("Year")+
  ylab("CO2 Emissions (metric tons per capita)")+
  ggtitle("China")
eur <- ts(data1$`European Union`, start = "1960")
autoplot(as.zoo(diff(eur)), geom = "line") + scale_x_yearqtr(format = "%Y")+
  theme_classic()+
  xlab("Year")+
  ylab("CO2 Emissions (metric tons per capita)")+
  ggtitle("European Union")

ind <- ts(data1$India, start="1960")
autoplot(as.zoo(diff(ind)), geom = "line") + scale_x_yearqtr(format = "%Y")+
  theme_classic()+
  xlab("Year")+
  ylab("CO2 Emissions (metric tons per capita)")+
  ggtitle("India")

us <- ts(data1$`United States`, start="1960")
autoplot(as.zoo(diff(us)), geom = "line") + scale_x_yearqtr(format = "%Y")+
  theme_classic()+
  xlab("Year")+
  ylab("CO2 Emissions (metric tons per capita)")+
  ggtitle("United States")
```

```{r newacfs,fig.show="hold", out.width="50%"}
acf(diff(br))
acf(diff(ch))
acf(diff(eur))
acf(diff(ind))
acf(diff(us))
```

\newpage
