---
#
# Required Information --------------------------------------------------------
# change lines 5 - 15
title: "Time Series Analysis"
type: "Final Project"
author: "Daniel Bustillo Mac Lean"
matriculation_number: "609728"
date: "31-08-2021"
university: "Humboldt-Universität zu Berlin"
institute: "School of Business and Economics" 
chair: "Institute for Statistics and Econometrics"
advisor1: "Prof. Dr. Simone Maxand"
program: "Master of Science in Economics"
#
# Optional Information --------------------------------------------------------
# comment out if unnecessary
chair2: "Chair of Econometrics"
abstract: |  
  `r if(knitr:::is_latex_output()) paste(readLines("prelim/00-abstract.Rmd"), collapse = '\n  ')`
acknowledge: |  
  `r if(knitr:::is_latex_output()) paste(readLines("prelim/00-acknowledge.Rmd"), collapse = '\n  ')`
abbreviations: |  
  `r if(knitr:::is_latex_output()) paste(readLines("prelim/00-abbreviations.Rmd"), collapse = '\n  ')`
lot: true                              # list of tables
lof: true                              # list of figures
link-citations: true
bibliography: bib/references.bib       # bibliography file
#
# Rendering Options -----------------------------------------------------------
#
knit: "bookdown::render_book"
output: 
  huwiwidown::thesis_pdf  
space_between_paragraphs: true       # if you want spaces bt paragraph
# header-includes:     # if including additional latex packages or commands
  # - \setlength{\parindent}{17pt}
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of
metadata used to produce the document.  Be careful with spacing in this header!
If you'd like to include a comment that won't be produced in your resulting file
enclose it in a block like this.
-->

<!--
If you receive a duplicate label error after knitting, make sure to delete the
index.Rmd file and then knit again.
-->

```{r include_packages, include = FALSE}
# This chunk ensures that the `huwiwidown` package is installed and loaded. This
# package includes the template files for the thesis.
if (!require(devtools)) {
  install.packages("devtools", repos = "http://cran.rstudio.com")
}
if (!require(huwiwidown)) {
  devtools::install_github("phister/huwiwidown")
  library(huwiwidown)
}
```

<!-- 
The {.unnumbered} option here means that the introduction will be "Section  0."
You can also use {-} for no numbers on section. 
-->

<!--
The body of the thesis starts underneath the text below. Feel free to delete
everything below this.
-->

# About this Thesis Template {-}

Welcome to the thesis template. This template is based on (and in many places 
copied directly from) the HU Berlin School of Business and Economics LaTeX 
template, but hopefully it will provide a nicer interface for those that have
never used LaTeX before. 

## Why use R Markdown? {-}

_R Markdown_ creates a simple and straightforward way to interface with the
beauty of LaTeX. Packages have been written in **R** to work directly with LaTeX
to produce nicely formatting tables and paragraphs. 

_R Markdown_ also allows you to read in your data, to analyze it and to
visualize it using **R** functions, and also to provide the documentation and
commentary on the results of your project. Using _R Markdown_ will also allow
you to easily keep track of your analyses in **R** chunks of code, with the 
resulting plots and output included as well.

<!--chapter:end:index.Rmd-->

# Introduction

* What is the subject of the study? Describe the economic/econometric problem.

* What is the purpose of the study (working hypothesis)?

* What do we already know about the subject (literature review)? Use citations:
  @lingenfelser2011systematic shows that... Alternative Forms of the Wald test are
  considered [@kuncheva2004combining].
  
* What is the innovation of the study?

* Provide an overview of your results.

* Outline of the paper:\
  _The paper is organized as follows. The next section describes the model
  under investigation. Section "[Data]" describes the data set
  and Section "[Results]" presents the results. Finally, Section
  "[Conclusion]" concludes._
* The introduction should not be longer than 4 pages.

<!--chapter:end:sections/01-intro.Rmd-->

# Methodology

* How was the data analyzed ?

* Present the underlying economic model/theory and give reasons why it is
  suitable to answer the given problem[^1].

* Present econometric/statistical estimation method and give reasons why it is
  suitable to answer the given problem.

* Allows the reader to judge the validity of the study and its findings.

* Depending on the topic this section can also be split up into separate
  sections.

[^1]: Here is an example of a footnote.

<!--chapter:end:sections/02-method.Rmd-->

# Data

* Describe the data and its quality.

* How was the data sample selected?

* Provide descriptive statistics such as:

  * time period,

  * item number of observations, data frequency,

  * item mean, median,

  * item min, max, standard deviation,

  * item skewness, kurtosis, Jarque--Bera statistic,

  * item time series plots, histogram.

* For example:

```{r table1, echo=FALSE}
col_names <- c(
  "3m", "6m", "1yr", "2yr", "3yr", "5yr", "7yr", "10yr", "12yr", "15yr"
)
means     <- c(
  3.138, 3.191, 3.307, 3.544, 3.756, 4.093, 4.354, 4.621, 4.741, 4.878
)
stddev    <- c(
  0.915, 0.919, 0.935, 0.910, 0.876, 0.825, 0.803, 0.776, 0.768, 0.762
)
row_names <- c("Mean", "StD")

df <- matrix(data = c(means, stddev), nrow = 2, byrow = T)
rownames(df) <- row_names
colnames(df) <- col_names
df <- data.frame(df)
knitr::kable(
  df,
  booktabs = TRUE,
  caption = "Detailed descriptive statistics of location and dispersion for 2100 observed swap rates for the period from February 15, 1999 to March 2, 2007. Swap rates measured as 3.12 (instead of 0.0312).",
  col.names = col_names, 
  escape = FALSE
) %>%
kable_styling(latex_options = c("HOLD_position")) %>%
row_spec(2, hline_after = T)
```

* Allows the reader to judge whether the sample is biased or to evaluate 
  possible impacts of outliers, for example.

* Here tables can be easily integrated using the `kable()` function in the
  `knitr` package (with perhaps some additional help from the `kableExtra`
  package). `kable()` will automatically generate a label for the table
  environment. That way you don't have to manually enter in the table in LaTex,
  you can embed tables from R code. 

* Tables can be referenced using `\@ref(label)`, where `label` is `tab:<name>`,
  where `<name>` is the code chunk label.

* The appearance may look different to tables directly typed with LaTex, due to
  limitations in `kable()`. To compare:
    \begin{table}[ht]

    \begin{center}
        {\footnotesize
        \begin{tabular}{l|cccccccccc}
            \hline \hline
                      & 3m    & 6m    & 1yr   & 2yr   & 3yr   & 5yr   & 7yr   & 10yr  & 12yr  & 15yr   \\
            \hline
                Mean   & 3.138 & 3.191 & 3.307 & 3.544 & 3.756 & 4.093 & 4.354 & 4.621 & 4.741 & 4.878  \\
                StD    & 0.915 & 0.919 & 0.935 & 0.910 & 0.876 & 0.825 & 0.803 & 0.776 & 0.768 & 0.762  \\
            \hline \hline
        \end{tabular}}
    \end{center}
    \caption{This table was handwritten with LaTeX.}
    \label{tab:table2}
    \end{table}

<!--chapter:end:sections/03-data.Rmd-->

# Results

* Organize material and present results.

* Use tables, figures (but prefer visual presentation):

  + Tables and figures should supplement (and not duplicate) the text.

  + Tables and figures should be provided with legends.

  + _Figure \@ref(fig:graph) shows how to include and reference graphics.
    The graphic must be labelled before. Files must be in __.eps__ format. You
    can do this really easily in R Markdown with `knitr::include_graphics()`_!

  + Figures can be referenced with `\@ref(fig:<name>)`, where `<name>` is the
    name of the code chunk.

```{r graph, echo=FALSE, fig.cap="Estimated residuals from model XXX. ...", out.width = "50%"}
knitr::include_graphics("figures/graph.eps")
```
  + Tables and graphics may appear in the text or in the appendix, especially if
  there are many simulation results tabulated, but is also depends on the study
  and number of tables resp. figures. The key graphs and tables must appear in
  the text!

* R Markdown can also supports math equations just like _LaTeX_!

  + _Equation \@ref(eq:SpecDens) represents the ACs of a stationary
    stochastic process:_
    \begin{equation}
            f_y(\lambda) = (2\pi)^{-1} \sum_{j=-\infty}^{\infty}
                           \gamma_j e^{-i\lambda j}
                         =(2\pi)^{-1}\left(\gamma_0 + 2 \sum_{j=1}^{\infty}
        \gamma_j \cos(\lambda j)\right)
                                       (\#eq:SpecDens)
    \end{equation}
    _where $i=\sqrt{-1}$ is the imaginary unit, $\lambda \in [-\pi, \pi]$ is the
    frequency and the $\gamma_j$ are the autocovariances of $y_t$._

  + Equations can be referenced with `\@ref(eq:<name>)`, where name is defined
    by adding `(\#eq:<name>)` in the line immediately before `\end{equation}`.

## Review of Results

*  Do the results support or do they contradict economic theory ?

*  What does the reader learn from the results?

*  Try to give an intuition for your results.

*  Provide robustness checks.

*  Compare to previous research.

<!--chapter:end:sections/04-results.Rmd-->

# Conclusion

* Give a short summary of what has been done and what has been found.

* Expose results concisely.

* Draw conclusions about the problem studied. What are the implications of your
  findings?

* Point out some limitations of study (assist reader in judging validity of
  findings).

* Suggest issues for future research.


<!--chapter:end:sections/05-conclusion.Rmd-->

\newpage

# References {-}

<!--
To remove the indentation of the first entry.
-->
\noindent

<!--
To create a hanging indent and spacing between entries.  These three lines may need to be removed for styles that don't require the hanging indent.
-->

\setlength{\parindent}{-0.5cm}
\setlength{\leftskip}{0.5cm}
\setlength{\parskip}{8pt}

<div id="refs"></div>

<!--
Restore to defaults
-->

\indent
\setlength{\parindent}{17pt}
\setlength{\leftskip}{0pt}
\setlength{\parskip}{0pt}

\newpage

<!--chapter:end:sections/98-references.Rmd-->

\appendix
# Appendix

Here goes the appendix!

## Figures

```{r graph2, echo=FALSE, out.extra = "", fig.cap="Estimated residuals (2) from model XXX. ...", out.width = "50%"}
knitr::include_graphics("figures/graph.eps")
```

## Tables

\begin{table}[ht]
    \begin{center}
        {\footnotesize
        \begin{tabular}{l|cccccccccc}
        \hline \hline
                        & 3m    & 6m    & 1yr   & 2yr   & 3yr   & 5yr   & 7yr   & 10yr  & 12yr  & 15yr   \\
            \hline
                Mean   & 3.138 & 3.191 & 3.307 & 3.544 & 3.756 & 4.093 & 4.354 & 4.621 & 4.741 & 4.878  \\
                Median & 3.013 & 3.109 & 3.228 & 3.490 & 3.680 & 3.906 & 4.117 & 4.420 & 4.575 & 4.759  \\
                Min    & 1.984 & 1.950 & 1.956 & 2.010 & 2.240 & 2.615 & 2.850 & 3.120 & 3.250 & 3.395  \\
                Max    & 5.211 & 5.274 & 5.415 & 5.583 & 5.698 & 5.805 & 5.900 & 6.031 & 6.150 & 6.295  \\
                StD    & 0.915 & 0.919 & 0.935 & 0.910 & 0.876 & 0.825 & 0.803 & 0.776 & 0.768 & 0.762  \\
            \hline \hline
        \end{tabular}}
    \end{center}
    \caption{Detailed descriptive statistics of location and dispersion for
    2100 observed swap rates for the period from
    February 15, 1999 to March 2, 2007. Swap rates measured as 3.12 (instead of 0.0312).}
    \label{tab:apptable}
\end{table}

\newpage

<!--chapter:end:sections/99-appendix.Rmd-->

